---
title: "Public health data Nepal, 2016"
format: 
  html: 
    page-layout: custom
execute:
  echo: true
editor: visual
server: shiny
---

```{r}
#| echo: false
#| warning: false
library(dplyr)
library(sf)
library(ggplot2)
library(shiny)
library(knitr)
```

## Maps related to public health at local levels in Nepal (2017/18)

## Loading the dataset

-   Health data: Count of various diseases

-   Health facilities data: Number of public and private health facilities in the local levels

    These datasets can be downloaded from the [~~*department of health services website*~~](https://dohs.gov.np/dohswebsite/public/ihims-raw-data) (link broken now) for the year 2017/18 AD (2074/75 BS)

-   Boundary of local levels: Geometry file (.shp) containing the boundary of local levels

    The boundary data is downloaded from [National geoportal of Nepal.](https://nationalgeoportal.gov.np/#/metadata/97)

## Number of health facilities

```{r}
#| echo: false
#| warning: false

health_facilities <- read.csv("data/raw_data/health_facilities.csv")
boundary <- read_sf("data/boundary/NepalLocalUnits0.shp")
boundary |> left_join(health_facilities, by = c("fid" ="fid")) -> joined_local_level

ggplot(data = joined_local_level) +
    geom_sf(aes(fill = Total.health.facilities..Public.and.Non.Public., geometry = geometry), 
            linewidth = 0.1, alpha = 0.5) +
    scale_fill_viridis_c(option = "D", trans = "sqrt") +
    theme_minimal() +
    theme(legend.position = "bottom") +
    labs(fill = "No of health facilities")

```

### Province-wise number of health facilities

```{r}
#| echo: false
#| panel: sidebar
# Create a dropdown for selecting provinces
provinces <- unique(joined_local_level$Provinces)
selectInput('selectedProvince', 'Select Province', provinces)
```

```{r}
#| panel: fill
#| echo: false
# Plot output for displaying the map
plotOutput('healthMap')
```

```{r}
#| context: server
library(dplyr)
library(sf)
library(ggplot2)
library(shiny)
health_count <- read.csv("data/raw_data/health_data.csv")
health_facilities <- read.csv("data/raw_data/health_facilities.csv")
boundary <- read_sf("data/boundary/NepalLocalUnits0.shp")
boundary |> left_join(health_facilities, by = c("fid" ="fid")) -> joined_local_level

selectedData <- reactive({
    joined_local_level |> filter(Provinces == input$selectedProvince)
  })
# Render the plot based on the selected province
output$healthMap <- renderPlot({
  data <- selectedData()
  # Generate the plot
  ggplot(data = data) +
    geom_sf(aes(fill = Total.health.facilities..Public.and.Non.Public., geometry = geometry), 
            linewidth = 0.1, alpha = 0.5) +
    scale_fill_viridis_c(option = "D", trans = "sqrt") +
    theme_minimal() +
    theme(legend.position = "bottom") +
    labs(fill = "No of health facilities")
})
```

## Total Outpatient Department (OPD) visitors

Showing local levels with OPD visitors less than 100,000 (16 local levels missing shown below in table)

```{r}
#| echo: false
#| warning: false
health_count <- read.csv("data/raw_data/health_data.csv")
health_count$total_opd <- health_count$Total.New.OPD.Visits.Male + health_count$Total.New.OPD.Visits.Female
boundary <- read_sf("data/boundary/NepalLocalUnits0.shp")
boundary |> left_join(health_count, by = c("fid" ="fid")) -> joined_local_level_1

ggplot(data = joined_local_level_1) +
  geom_sf(aes(fill = ifelse(total_opd>100000, NA, total_opd),
              geometry = geometry),
          linewidth = 0.1, alpha = 0.5) +
  scale_fill_viridis_c(option = "D", trans = "sqrt", na.value = "transparent") +
  theme_minimal() +
  theme(legend.position = "bottom",
        legend.key.width = unit(2, "cm")) +
  labs(fill = "No of OPD visitors")

```

The local levels with high number of OPD visitors are following 16 local levels which are mostly large cities (also related to presence of many health facilities):

```{r}
#| echo: false
#| warning: false
# Selecting required columns: Local Level Name, District Name, and Malaria case count
selected_data <- joined_local_level_1 |>
  select(Name.of.Government.Unit.x, DISTRICT, Provinces.y, total_opd) |>
  arrange(desc(total_opd)) |> st_drop_geometry()

# Display the top 10
top_10_local_levels <- head(selected_data, 16)

kable(top_10_local_levels, 
      caption = "Local levels with more than 100,000 OPD visitors",
      col.names = c("Local Level", "District", "Province", "OPD visitors"),
      align = c("l", "l", "l", "r"))
```

```         
"Total.Diarrhoea.cases", "Fully.Immunized.Children" 
```

# Infectious disease

## Total number of Malaria cases

```{r}
#| echo: false
#| warning: false


ggplot(data = joined_local_level_1) +
  geom_sf(aes(fill = ifelse(Malaria..Total.number.of.positive.cases == 0, 
                            NA, 
                            Malaria..Total.number.of.positive.cases),
              geometry = geometry),
          linewidth = 0.1, alpha = 0.5) +
  scale_fill_viridis_c(option = "D", trans = "sqrt", na.value = "transparent") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(fill = "No of Malaria cases")
```

It is worth noting that, although malaria cases are primarily clustered in the southern regions (the plains of the Terai), there are notable exceptions where high case numbers are reported in the hilly and mountainous districts, particularly in the western region. Here is a list of several notable local levels that indicate a higher distribution of cases in the Western and Far Western regions.

```{r}
#| echo: false
#| warning: false
# Selecting required columns: Local Level Name, District Name, and Malaria case count
selected_data <- joined_local_level_1 |>
  select(Name.of.Government.Unit.x, DISTRICT, Provinces.y, Malaria..Total.number.of.positive.cases) |>
  arrange(desc(Malaria..Total.number.of.positive.cases)) |> st_drop_geometry()

# Display the top 10
top_10_local_levels <- head(selected_data, 10)

kable(top_10_local_levels, 
      caption = "Top 10 Local Levels with Malaria Cases",
      col.names = c("Local Level", "District", "Province", "Malaria Total.number.of.positive.cases"),
      align = c("l", "l", "l", "r"))
```

## Total number of Tuberculosis cases

```{r}
#| echo: false
#| warning: false


ggplot(data = joined_local_level_1) +
  geom_sf(aes(fill = TB.PBC..new..registered.cases,
              geometry = geometry),
          linewidth = 0.1, alpha = 0.5) +
  scale_fill_viridis_c(option = "D", trans = "sqrt", na.value = "transparent") +
  theme_minimal() +
  theme(legend.position = "bottom",
        legend.key.width = unit(1.5, "cm")) +
  labs(fill = "No of Tuberculosis  cases")
```

The cases of Tuberculosis are mostly clustered in bigger cities as listed below:

```{r}
#| echo: false
#| warning: false
# Selecting required columns: Local Level Name, District Name, and Malaria case count
selected_data <- joined_local_level_1 |>
  select(Name.of.Government.Unit.x, DISTRICT, Provinces.y, TB.PBC..new..registered.cases) |>
  arrange(desc(TB.PBC..new..registered.cases)) |> st_drop_geometry()

# Display the top 10
top_10_local_levels <- head(selected_data, 20)

kable(top_10_local_levels, 
      caption = "Top 10 Local Levels with Malaria Cases",
      col.names = c("Local Level", "District", "Province ", "New TB cases"),
      align = c("l", "l", "l", "r"))
```

Here we look at the historical TB patient's information in the number of cured TB cases

```{r}
#| echo: false
#| warning: false


ggplot(data = joined_local_level_1) +
  geom_sf(aes(fill = TB.PBC..new..treatment.completed....cured.and.completed..cases,
              geometry = geometry),
          linewidth = 0.1, alpha = 0.5) +
  scale_fill_viridis_c(option = "D", trans = "sqrt", na.value = "transparent") +
  theme_minimal() +
  theme(legend.position = "bottom",
        legend.key.width = unit(1.5, "cm")) +
  labs(fill = "No of Cured Tuberculosis  cases")
```

```{r}
#| echo: false
#| warning: false
# Selecting required columns: Local Level Name, District Name, and Malaria case count
selected_data <- joined_local_level_1 |>
  select(Name.of.Government.Unit.x, DISTRICT, Provinces.y, TB.PBC..new..treatment.completed....cured.and.completed..cases) |>
  arrange(desc(TB.PBC..new..treatment.completed....cured.and.completed..cases)) |> st_drop_geometry()

# Display the top 10
top_10_local_levels <- head(selected_data, 20)

kable(top_10_local_levels, 
      caption = "Top 10 Local Levels with Malaria Cases",
      col.names = c("Local Level", "District", "Province ", "Cured TB cases"),
      align = c("l", "l", "l", "r"))
```
