---
title-block-banner: true
title: "Public health data Nepal, 2016"
date: 2025-02-01
author-title: Biplov Parajuli and Poshan Niraula
format: html
editor: visual
---

```{r}
#| echo: false
#| warning: false
library(dplyr)
library(sf)
library(ggplot2)
library(shiny)
library(knitr)
```

## Maps related to public health at local levels in Nepal (2017/18)

## Loading the dataset

-   Health data: Count of various diseases

-   Health facilities data: Number of public and private health facilities in the local levels

    These datasets can be downloaded from the [~~*department of health services website*~~](https://dohs.gov.np/dohswebsite/public/ihims-raw-data) (link broken now) for the year 2017/18 AD (2074/75 BS)

-   Boundary of local levels: Geometry file (.shp) containing the boundary of local levels

    The boundary data is downloaded from [National geoportal of Nepal.](https://nationalgeoportal.gov.np/#/metadata/97)

## Number of health facilities

```{r}
#| echo: false
#| warning: false

health_facilities <- read.csv("data/raw_data/health_facilities.csv")
boundary <- read_sf("data/boundary/NepalLocalUnits0.shp")
boundary |> left_join(health_facilities, by = c("fid" ="fid")) -> joined_local_level

ggplot(data = joined_local_level) +
    geom_sf(aes(fill = Total.health.facilities..Public.and.Non.Public., geometry = geometry), 
            linewidth = 0.1, alpha = 0.5) +
    scale_fill_viridis_c(option = "D", trans = "sqrt") +
    theme_minimal() +
    theme(legend.position = "bottom") +
    labs(fill = "No of health facilities")

```

## Total Outpatient Department (OPD) visitors

Showing local levels with OPD visitors less than 100,000 (16 local levels missing shown below in table)

```{r}
#| echo: false
#| warning: false
health_count <- read.csv("data/raw_data/health_data.csv")
health_count$total_opd <- health_count$Total.New.OPD.Visits.Male + health_count$Total.New.OPD.Visits.Female
boundary <- read_sf("data/boundary/NepalLocalUnits0.shp")
boundary |> left_join(health_count, by = c("fid" ="fid")) -> joined_local_level_1

ggplot(data = joined_local_level_1) +
  geom_sf(aes(fill = ifelse(total_opd>100000, NA, total_opd),
              geometry = geometry),
          linewidth = 0.1, alpha = 0.5) +
  scale_fill_viridis_c(option = "D", trans = "sqrt", na.value = "transparent") +
  theme_minimal() +
  theme(legend.position = "bottom",
        legend.key.width = unit(2, "cm")) +
  labs(fill = "No of OPD visitors")

```

The local levels with high number of OPD visitors are following 16 local levels which are mostly large cities (also related to presence of many health facilities):

```{r}
#| echo: false
#| warning: false
# Selecting required columns: Local Level Name, District Name, and Malaria case count
selected_data <- joined_local_level_1 |>
  select(Name.of.Government.Unit.x, DISTRICT, Provinces.y, total_opd) |>
  arrange(desc(total_opd)) |> st_drop_geometry()

# Display the top 10
top_10_local_levels <- head(selected_data, 16)

kable(top_10_local_levels, 
      caption = "Local levels with more than 100,000 OPD visitors",
      col.names = c("Local Level", "District", "Province", "OPD visitors"),
      align = c("l", "l", "l", "r"))
```

```         
"Total.Diarrhoea.cases", "Fully.Immunized.Children" 
```

# Infectious disease

## Total number of Malaria cases

```{r}
#| echo: false
#| warning: false
library(ggforce)

circle_data <- data.frame(
  x = 81.615917, # Longitude of the center
  y = 29.464046, # Latitude of the center
  r = 0.5   # Radius of the circle
)

ggplot(data = joined_local_level_1) +
  geom_sf(aes(fill = ifelse(Malaria..Total.number.of.positive.cases == 0, 
                            NA, 
                            Malaria..Total.number.of.positive.cases),
              geometry = geometry),
          linewidth = 0.1, alpha = 0.5) +
    geom_circle(data = circle_data,
               aes(x0 = x, y0 = y, r = r),
               color = "red", fill = NA, size = 1) +
  coord_sf(datum = NA) +
  scale_fill_viridis_c(option = "D", trans = "sqrt", na.value = "transparent") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  labs(fill = "No of Malaria cases")
```

It is worth noting that, although malaria cases are primarily clustered in the southern regions (the plains of the Terai = hot and humid climate), there are notable exceptions where high case numbers are reported in the hilly and mountainous districts, particularly in the western region. Here is a list of several notable local levels that indicate a higher distribution of cases in the Western and Far Western regions.

```{r}
#| echo: false
#| warning: false

library(ggforce)
# Selecting required columns: Local Level Name, District Name, and Malaria case count
selected_data <- joined_local_level_1 |>
  select(Name.of.Government.Unit.x, DISTRICT, Provinces.y, Malaria..Total.number.of.positive.cases) |>
  arrange(desc(Malaria..Total.number.of.positive.cases)) |> st_drop_geometry()

# Display the top 10
top_10_local_levels <- head(selected_data, 10)

kable(top_10_local_levels, 
      caption = "Top 10 Local Levels with Malaria Cases",
      col.names = c("Local Level", "District", "Province", "Malaria Total.number.of.positive.cases"),
      align = c("l", "l", "l", "r"))
```

## Total number of Tuberculosis cases

```{r}
#| echo: false
#| warning: false



ggplot(data = joined_local_level_1) +
  geom_sf(aes(fill = TB.PBC..new..registered.cases,
              geometry = geometry),
          linewidth = 0.1, alpha = 0.5) +
  scale_fill_viridis_c(option = "D", trans = "sqrt", na.value = "transparent") +
  theme_minimal() +
  theme(legend.position = "bottom",
        legend.key.width = unit(1.5, "cm")) +
  labs(fill = "No of Tuberculosis  cases")
```

The cases of Tuberculosis are mostly clustered in bigger cities as listed below:

```{r}
#| echo: false
#| warning: false
# Selecting required columns: Local Level Name, District Name, and Malaria case count
selected_data <- joined_local_level_1 |>
  select(Name.of.Government.Unit.x, DISTRICT, Provinces.y, TB.PBC..new..registered.cases) |>
  arrange(desc(TB.PBC..new..registered.cases)) |> st_drop_geometry()

# Display the top 10
top_10_local_levels <- head(selected_data, 20)

kable(top_10_local_levels, 
      caption = "Top 10 Local Levels with Tuberculosis Cases",
      col.names = c("Local Level", "District", "Province ", "New TB cases"),
      align = c("l", "l", "l", "r"))
```

Here we look at the historical TB patient's information in the number of cured TB cases

```{r}
#| echo: false
#| warning: false


ggplot(data = joined_local_level_1) +
  geom_sf(aes(fill = TB.PBC..new..treatment.completed....cured.and.completed..cases,
              geometry = geometry),
          linewidth = 0.1, alpha = 0.5) +
  scale_fill_viridis_c(option = "D", trans = "sqrt", na.value = "transparent") +
  theme_minimal() +
  theme(legend.position = "bottom",
        legend.key.width = unit(1.5, "cm")) +
  labs(fill = "No of Cured Tuberculosis  cases")
```

```{r}
#| echo: false
#| warning: false
# Selecting required columns: Local Level Name, District Name, and Malaria case count
selected_data <- joined_local_level_1 |>
  select(Name.of.Government.Unit.x, DISTRICT, Provinces.y, TB.PBC..new..treatment.completed....cured.and.completed..cases) |>
  arrange(desc(TB.PBC..new..treatment.completed....cured.and.completed..cases)) |> st_drop_geometry()

# Display the top 10
top_10_local_levels <- head(selected_data, 20)

kable(top_10_local_levels, 
      caption = "Top 10 Local Levels with cureed Tuberculosis Cases",
      col.names = c("Local Level", "District", "Province ", "Cured TB cases"),
      align = c("l", "l", "l", "r"))
```
